On the Dynamics of Anthropomorphism in Robotics \thanks{This research was supported by the Swiss National Science Foundation through the National Centre of Competence in Research Robotics.}

### 2013-05-22 anthropomorphism.tex
Work on the 'cognitive model of anthropomorphism' section

	"[...] the experiment demonstrates that human beings implicitly attribute human-like qualities, such as mental states, to nonhuman agents. This finding was evident on a neurophysiological level as well as behavioural level." \cite{hegel_understanding_2008}

@@

	"On a next level, \textbf{stereotypical behaviour}, which expresses e.g. shyness of curiosity, addresses more complex processes but adds essentially to anthropomorphic strategies [38] [46]." \cite{schmitz_concepts_2011}

@@

The second phase, \emph{familiarization}, lasts longer (up to several days) and
models the process of the human getting acquainted to the robot: by observation
and interaction, the human builds a model of the robot's behaviour that allows
him/her to predict the robot's actions. We observe a decrease of
anthropomorphic effects during this phase, that we explain by the acquired
ability to predict the behaviour of the robot: the initial apparent behavioural
complexity vanishes, and the robot is considered more and more as a machine.
Section~\ref{sec:familiarization} discusses the second phase.

@@

\paragraph{Disruptive Behaviours}

@@

@@

\draw (2.2,1) node[align=left] {\tiny{\it observation (shape, motion, sound)}}; %label
\draw (5.2,3) node[align=left] {\tiny \it observation (interactive behaviour) \\ \tiny \it or short interaction}; %label
\draw (7.8,5) node[align=left] {\tiny \it contextualized interaction}; %label

@@

@@

We emit the hypothesis that an unexpected robot behavior leads the user to
suddently restate its behavioral model of the robot and will temporarly lead to
an increase of the level of anthropomorphism (see the spikes on
Figure~\ref{fig:dynamics}).

@@

%% experience -> unexpected behaviour should not impact the interaction child-robot BEFORE the theory of mind is established!

>>>>

	"[...] the experiment demonstrates that human beings implicitly attribute human-like qualities, such as mental states, to nonhuman agents. This finding was evident on a neurophysiological level as well as behavioral level." \cite{hegel_understanding_2008}

@@

	"On a next level, \textbf{stereotypical behavior}, which expresses e.g. shyness of curiosity, addresses more complex processes but adds essentially to anthropomorphic strategies [38] [46]." \cite{schmitz_concepts_2011}

@@

The second phase, \emph{familiarization}, lasts longer (up to several days) and
models the process of the human getting acquainted to the robot: by observation
and interaction, the human builds a model of the robot's behavior that allows
him/her to predict the robot's actions. We observe a decrease of
anthropomorphic effects during this phase, that we explain by the acquired
ability to predict the behavior of the robot: the initial apparent behavioral
complexity vanishes, and the robot is considered more and more as a machine.
Section~\ref{sec:familiarization} discusses the second phase.

@@

\paragraph{Disruptive behaviors}
\ref{sec:disruptive}

\paragraph{Expectation mismatch}

@@

\fxnote{Note that ICA and SLA are NOT correlated}

@@

\draw (2.2,1) node[align=left] {\tiny{\it observation (shape, motion, sound)}}; %label
\draw (5.2,3) node[align=left] {\tiny \it observation (interactive behavior) \\ \tiny \it or short interaction}; %label
\draw (7.8,5) node[align=left] {\tiny \it contextualized interaction}; %label

@@

The first cognitive phase we identify that contribute to anthropomorphism is
actually \emph{pre-cognitive}.\fxfatal{Complete that. Reference to the pleo
paper}.

After a longer observation period (typically including complete action
sequences of the robot) or short interaction (touching, short talk like
greetings), the human enters the cognitive phase II: in this phase, the human
starts to build a behavioral and cognitive model of the robot that would
support both the observed and imagined capabilities of the robot. Based on the
\emph{familiarity thesis}\fxfatal{citation!}, we suggest that the human first
projects onto the robot mental models of similar agents he/she is already
familiar with. This may range from animals to adults, through pets and
children. We hypothesise that the nature of the projected mental model, as well
as how deep the human engage in this projection, are driven by the same
parameters as we presented for the \emph{initial capital of anthropomorphism}
(section~\ref{sec:ica}, p.\pageref{sec:ica}).

We identify a further transition to the cognitive phase III after a
\emph{contextualized} interaction. A \emph{contextualized} interaction is
\emph{explicitly purposeful} (the purpose of the interaction, be it purely
entertainment, is explicit and conscious for the human), and takes place in an
environment that fosters a stronger cognitive (and possibly affective)
commitment from the human in the interaction (typically, at home). During this
interaction, the human iteratively restate and reshape its behavioral and
mental model of the robot (\emph{how does the robot react to such and such
situation/input? what does the robot knows about me? about itself? about our
environment? what can the robot learn?}, etc.).

This mental process heavily depends on the human understanding of the robot
inner working, as well as his/her own tendency to anthropomorphize (the
\emph{personality} in ICA factor), but at this stage, the \emph{perception} of
the robot (its shape for instance) and its intended \emph{purpose} play only a
minor role. It is mostly a human-centric process.

\paragraph{Relation to the model of anthropomorphism} It must be noted that the
three cognitive phases we introduce here do not match the
\emph{initialization}, \emph{familiarization} and \emph{stabilization} phases
introduced at~\ref{sec:phases}: in particular, cognitive phases I and II are
both included in the \emph{initialization} phase of the anthropomorphism model.

Sub-cognitive anthropomorphism typically \emph{initiates} the novelty effect by
rapidly engaging the human in the interaction through this initial projected
\emph{agency}, whereas cognitive phase II (projection of familiar mental
models) supports the novelty effect by inducing beliefs that the robot is set
up with possibly complex cognitive abilities.

The cognitive phase II also overlaps with the \emph{Familiarization} phase: as
(s)he get used to the robot, we hypothesize the human restate and adapt its
cognitive model of the robot by iteratively reshaping pre-existant, familiar
models until it provides a satisfying support to explain and justify the
observed robot behavior.

A \emph{stable level of anthropomorphism} is reached when the adaptation
process depicted in cognitive phase III reached a stable state, \ie the human
experience with the robot is correctly supported by the cognitive model the
human has built.

\paragraph{Limits} This discussion on the cognitive correlates of the dynamics
of anthropomorphism are speculative, and only indirectly supported by
experimental evidence. New experiences need to be designed to specifically test
these hypothesis.\fxnote{...}

@@

We emit the hypothesis that an unexpected robot behavior leads the user to
suddently restate its behavioral model of the robot and will temporarly lead to
an increase of the level of anthropomorphism (depicted by the spikes on
Figure~\ref{fig:dynamics}).

@@

%% experience -> unexpected behavior should not impact the interaction child-robot BEFORE the theory of mind is established!
